---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(tidyverse)
library(sf)
library(geotidy)
```
# geotidy

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/etiennebr/geotidy.svg?branch=master)](https://travis-ci.org/etiennebr/geotidy)
[![Codecov test coverage](https://codecov.io/gh/etiennebr/geotidy/branch/master/graph/badge.svg)](https://codecov.io/gh/etiennebr/geotidy?branch=master)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

Manipulate spatial data tidily. `Geotidy` provides a selection of spatial
functions from the `sf` package that are adapted to a tidy workflow. It relies
on tibbles and `dplyr` verbs, and forces you to be explicit about the spatial
operations you desire. The package is experimental. If you want to learn more
about the motivations behind it, see the [Motivation] section.

## Installation

You can install the experimental version of geotidy from github :
```{r eval = FALSE}
# install.packages("remotes")
remotes::install_github("etiennebr/geotidy")
```

## Example

Here's how you can create spatial data.

```{r example, message = FALSE}
library(tidyverse)
library(geotidy)

tibble(place = "Sunset Room", longitude = -97.7404985, latitude = 30.2645315) %>% 
  mutate(geometry = st_point(longitude, latitude))
```



```{r}
library(tidyverse)
library(sf)
library(geotidy)  # note: load geotidy last, because it replaces some sf functions

giza <- tribble(
  ~what,         ~longitude, ~latitude,
  "Giza",        31.1342,   29.9792,
  "Khafre",      31.130833, 29.976111,
  "Menkaure",    31.128333, 29.9725,
  "Khentkaus I", 31.135608, 29.973406,
  "Sphynx",      31.137778, 29.975278
) %>% 
  # add a geometry column
  mutate(geometry = st_point(longitude, latitude) %>% st_set_crs(4326))
giza

giza %>% 
  ggplot(aes(geometry = geometry, label = what)) + 
  geom_sf(size = 10, alpha = 0.1) +
  geom_sf_text()

# create a buffer around the points and union in a single geometry
giza %>% 
  mutate(
    projected = st_transform(geometry, 32636),
    buffer = st_buffer(projected, units::as_units(250, "m"))
  ) %>% 
  summarise(
    geometry = st_union(projected),
    buffer = st_union(buffer)
  ) %>% 
  ggplot(aes(geometry = buffer)) + 
  geom_sf()
```

## Motivation

Many spatial formats use a model where a geometry can have many attributes. This
model can be very useful in many applications, but it sometimes conflicts with
the principles of tidy data, where one observation is one row. To keep spatial
data and operations tidy, `geotidy` does less than `sf`. It makes manipulations
explicit by forcing the use of a function on the geometry column. It doesn't try
to guess which is the geometry column that should receive the operation. It also
makes it clear by reading the code, which geometry is impacted. This is done by
treating geometry columns just like other tibble columns. `sf` often hides the
geometry column, `geotidy` treats it just like a regular columns. This also
makes it easier to interact with other OGC compliant tools, such as `postgis` or
`spark+geomesa`.

If you already use `dplyr` with `sf`, `geotidy` should fell natural and remove
some of the casting operations. `geotidy` guarantees that your data will stay
tidy from start to finish. By having explicit management of geometry columns, it
is also easy to track multiple columns.

`geotidy` also guarantees that the returned values are either scalar, a
vector or a list with the same length than the original geometry and not drop
any data without the user explicit consent (looking at you `st_cast`!). 
We also make it clear which functions aggregate the data by using `group_by`.

`geotidy` is not a fork or a separation from `sf`. It just adds a constrained
layer on top of `sf` to facilitate a tidy workflow. It is an experiment that
could be integrated in `sf` and is likely to change.
